The landscape of chloroplast genome assembly tools

### 2019-06-06 bmc_article.tex
Update on Overleaf.

@@

Not all the tools were able to benefit from having access to multiple threads. Both \np{} and \oa{} take about the same time independent of being able to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads. \cassp{}, \ce{}, \go  and \fp{} all profit from multi-threading (\cref{fig:performance_runtime}, \cref{fig:performance_memory_cpu}).

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

\subsubsection*{Real data sets}
Concerning the performance of the assemblers on the real data sets, we were able to observe considerable differences in the median score (\cref{fig:swarmplot}).
The highest scores were achieved by \go{} with a median of \num{99.7} and \num{199} circular assemblies out of a total of 356 assemblies that resulted in an output (\cref{tab:scores}).
The performance of \go{} is followed by \fp{}, \np{}, \oa{}, and \ce{}. \fp{} is outperforming the latter two slightly in terms of score, with twice as much \num{114} perfectly assembled chloroplast genomes (\np{} produced \num{66} and \oa{} \num{55} circular genomes).
\ioga{} and \cassp{} were both not able to assemble a circular, single-contig genome (\cref{tab:scores}), consequently resulting in the lowest mean and median scores (\cref{fig:upset}).

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{plots/real_datasets_tree.pdf}
  \caption{\csentence{Success of assemblers on real data sets on tree derived from NCBI taxonomy \cite{ncbi2011}}
       }
      \label{fig:tree}
      \end{figure}\todo{cite itol}

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Docker images used in our benchmark setup.}
\label{tab:dockerimages}
\centering
      \resizebox{\textwidth}{!}{\begin{tabular}{ccc}
        \hline
          Tool & Image name and tag & SHA256 Checksum   \\ \hline
          \ce & \dockerce & \dockercesha \\
          \cassp & \dockercassp & \dockercasspsha \\
          \fp & \dockerfp & \dockerfpsha \\
          \go & \dockergo & \dockergosha \\
          \ioga & \dockerioga & \dockeriogasha \\
          \np & \dockernp & \dockernpsha \\
          \oa & \dockeroa & \dockeroasha \\ \hline
      \end{tabular}}
\end{table}

\begin{table}[]
    \centering
    \caption{Overview of the results of the qualitative usability evaluation. Each tool 
    could score \good{}, \ok{} or \bad{} in each of the categories.}
    \label{tab:resultsqual}
\begin{tabular}{p{3cm}cccccc}   
Tool & Installation & Test/Tutorial & Documentation & Maintenance & FLOSS\\                           \hline \ce{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \good{} \\
\cassp{}  &  \ok{}    &  \good{}  &  \ok{}    &  \good{}  &  \good{} \\
\fp{}     &  \bad{}   &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\go{}     &  \ok{}    &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\ioga{}   &  \bad{}   &  \bad{}   &  \ok{}    &  \ok{}    &  \bad{}  \\
\np{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \ok{}   \\
\oa{}     &  \bad{}   &  \bad{}   &  \ok{}    &  \good{}  &  \good{} \\ \hline
\end{tabular}      
\end{table}

\begin{table}[ht]
\caption{Scores of assemblies of simulated data}
\label{tab:scores}
\centering
\begin{tabular}{rlrrrrrrr}
  \hline
 & data set & CAP & CE & Fast-Plast & GetOrganelle & IOGA & NOVOPlasty & org.ASM \\ 
  \hline
1 & sim\_150bp.0-1 & 79.10 & 100.00 & 99.72 & 100.00 &  & 91.52 & 100.00 \\ 
  2 & sim\_150bp.1-10 &  & 56.44 &  & 100.00 &  & 91.52 & 78.00 \\ 
  3 & sim\_150bp.1-10.2M &  &  & 99.97 & 100.00 &  & 100.00 & 82.72 \\ 
  4 & sim\_150bp.1-100 &  & 100.00 &  & 100.00 &  & 91.52 & 91.50 \\ 
  5 & sim\_150bp.1-100.2M &  & 100.00 & 99.47 & 100.00 & 74.82 & 100.00 & 100.00 \\ 
  6 & sim\_150bp.1-1000 & 79.10 &  &  & 100.00 &  & 91.52 & 91.50 \\ 
  7 & sim\_150bp.1-1000.2M & 79.10 & 100.00 & 99.72 & 100.00 &  & 91.52 & 100.00 \\ 
  8 & sim\_250bp.0-1 & 79.10 & 100.00 & 93.83 & 100.00 &  & 91.52 & 100.00 \\ 
  9 & sim\_250bp.1-10 &  & 54.98 & 68.45 & 100.00 &  & 91.52 & 40.20 \\ 
  10 & sim\_250bp.1-10.2M &  &  & 93.00 & 100.00 &  & 87.40 & 40.20 \\ 
  11 & sim\_250bp.1-100 &  & 100.00 & 93.83 & 100.00 & 65.81 & 91.52 & 100.00 \\ 
  12 & sim\_250bp.1-100.2M &  & 100.00 & 93.83 & 100.00 & 75.73 & 87.40 & 100.00 \\ 
  13 & sim\_250bp.1-1000 & 79.10 & 21.30 &  & 100.00 &  & 91.52 & 91.50 \\ 
  14 & sim\_250bp.1-1000.2M & 79.10 & 100.00 & 93.83 & 100.00 &  & 87.40 & 100.00 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Mean scores of chloroplast genome assemblers}
\label{tab:scores}
\centering
\begin{tabular}{rlrrrr}
  \hline
 & assembler & Mean & SD & N\_perfect & N\_tot \\ 
  \hline
  1 & CAP & 51.26 & 8.42 &   0 & 221 \\ 
  2 & CE & 69.99 & 12.33 &  14 & 205 \\ 
  3 & Fast-Plast & 86.96 & 15.33 & 114 & 352 \\ 
  4 & GetOrganelle & 89.13 & 15.30 & 199 & 356 \\ 
  5 & IOGA & 68.74 & 11.69 &   0 & 296 \\ 
  6 & NOVOPlasty & 82.78 & 16.55 &  66 & 270 \\ 
  7 & org.ASM & 82.77 & 16.38 &  55 & 228 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

Not all the tools were able to benefit from having access to multiple threads. Both \np{} and \oa{} take about the same time independent of being able to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads. \cassp{}, \ce{}, \go  and \fp{} all profit from multi-threading (\cref{fig:performance_runtime,fig:performance_memory_cpu}).

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated,tab:scores_simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

\subsubsection*{Real data sets}
Concerning the performance of the assemblers on the real data sets, we were able to observe considerable differences in the median score (\cref{fig:swarmplot}).
The highest scores were achieved by \go{} with a median of \num{99.7} and \num{199} circular assemblies out of a total of 356 assemblies that resulted in an output (\cref{tab:scores_real}).
The performance of \go{} is followed by \fp{}, \np{}, \oa{}, and \ce{}. \fp{} is outperforming the latter two slightly in terms of score, with twice as much \num{114} perfectly assembled chloroplast genomes (\np{} produced \num{66} and \oa{} \num{55} circular genomes).
\ioga{} and \cassp{} were both not able to assemble a circular, single-contig genome (\cref{tab:scores}), consequently resulting in the lowest mean and median scores (\cref{fig:upset}).

@@

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{plots/real_datasets_tree.pdf}
  \caption{\csentence{Success of assemblers on real data sets on tree derived from NCBI taxonomy \cite{ncbi2011}. Plot was prepared using \cite{Letunic2019}}
       }
      \label{fig:tree}
      \end{figure}

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[]
    \centering
    \caption{Overview of the results of the qualitative usability evaluation. Each tool 
    could score \good{}, \ok{} or \bad{} in each of the categories.}
    \label{tab:resultsQual}
\begin{tabular}{p{3cm}cccccc}   
Tool & Installation & Test/Tutorial & Documentation & Maintenance & FLOSS\\                           \hline \ce{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \good{} \\
\cassp{}  &  \ok{}    &  \good{}  &  \ok{}    &  \good{}  &  \good{} \\
\fp{}     &  \bad{}   &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\go{}     &  \ok{}    &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\ioga{}   &  \bad{}   &  \bad{}   &  \ok{}    &  \ok{}    &  \bad{}  \\
\np{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \ok{}   \\
\oa{}     &  \bad{}   &  \bad{}   &  \ok{}    &  \good{}  &  \good{} \\ \hline
\end{tabular}      
\end{table}

\begin{table}[ht]
\caption{Scores of assemblies of simulated data}
\label{tab:scores_simulated}
\centering
\begin{tabular}{rlrrrrrrr}
  \hline
 & data set & CAP & CE & Fast-Plast & GetOrganelle & IOGA & NOVOPlasty & org.ASM \\ 
  \hline
1 & sim\_150bp.0-1 & 79.10 & 100.00 & 99.72 & 100.00 &  & 91.52 & 100.00 \\ 
  2 & sim\_150bp.1-10 &  & 56.44 &  & 100.00 &  & 91.52 & 78.00 \\ 
  3 & sim\_150bp.1-10.2M &  &  & 99.97 & 100.00 &  & 100.00 & 82.72 \\ 
  4 & sim\_150bp.1-100 &  & 100.00 &  & 100.00 &  & 91.52 & 91.50 \\ 
  5 & sim\_150bp.1-100.2M &  & 100.00 & 99.47 & 100.00 & 74.82 & 100.00 & 100.00 \\ 
  6 & sim\_150bp.1-1000 & 79.10 &  &  & 100.00 &  & 91.52 & 91.50 \\ 
  7 & sim\_150bp.1-1000.2M & 79.10 & 100.00 & 99.72 & 100.00 &  & 91.52 & 100.00 \\ 
  8 & sim\_250bp.0-1 & 79.10 & 100.00 & 93.83 & 100.00 &  & 91.52 & 100.00 \\ 
  9 & sim\_250bp.1-10 &  & 54.98 & 68.45 & 100.00 &  & 91.52 & 40.20 \\ 
  10 & sim\_250bp.1-10.2M &  &  & 93.00 & 100.00 &  & 87.40 & 40.20 \\ 
  11 & sim\_250bp.1-100 &  & 100.00 & 93.83 & 100.00 & 65.81 & 91.52 & 100.00 \\ 
  12 & sim\_250bp.1-100.2M &  & 100.00 & 93.83 & 100.00 & 75.73 & 87.40 & 100.00 \\ 
  13 & sim\_250bp.1-1000 & 79.10 & 21.30 &  & 100.00 &  & 91.52 & 91.50 \\ 
  14 & sim\_250bp.1-1000.2M & 79.10 & 100.00 & 93.83 & 100.00 &  & 87.40 & 100.00 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[h!]
\caption{Docker images used in our benchmark setup.}
\label{tab:dockerimages}
\centering
      \resizebox{\textwidth}{!}{\begin{tabular}{ccc}
        \hline
          Tool & Image name and tag & SHA256 Checksum   \\ \hline
          \ce & \dockerce & \dockercesha \\
          \cassp & \dockercassp & \dockercasspsha \\
          \fp & \dockerfp & \dockerfpsha \\
          \go & \dockergo & \dockergosha \\
          \ioga & \dockerioga & \dockeriogasha \\
          \np & \dockernp & \dockernpsha \\
          \oa & \dockeroa & \dockeroasha \\ \hline
      \end{tabular}}
\end{table}

\begin{table}[ht]
\caption{Mean scores of chloroplast genome assemblers}
\label{tab:scores_real}
\centering
\begin{tabular}{rlrrrr}
  \hline
 & assembler & Mean & SD & N\_perfect & N\_tot \\ 
  \hline
  1 & CAP & 51.26 & 8.42 &   0 & 221 \\ 
  2 & CE & 69.99 & 12.33 &  14 & 205 \\ 
  3 & Fast-Plast & 86.96 & 15.33 & 114 & 352 \\ 
  4 & GetOrganelle & 89.13 & 15.30 & 199 & 356 \\ 
  5 & IOGA & 68.74 & 11.69 &   0 & 296 \\ 
  6 & NOVOPlasty & 82.78 & 16.55 &  66 & 270 \\ 
  7 & org.ASM & 82.77 & 16.38 &  55 & 228 \\ 
   \hline
\end{tabular}
\end{table}
